Binary companions of evolved stars in \apogee\ \DR: \\ Search method and catalog of $\sim$5,000 companions

### 2018-04-24 paper/1-catalog/twoface-catalog.tex
Fix MCMC parametrization
Address referee comments

The \apogee\ surveys primarily target red-giant-branch (\RGB) and other evolved
stars (e.g., red clump giants, RC), which are ideal for the study of single-line
binary systems.
In general, they are unlikely to have comparably-bright companions, and their
spectra are therefore fit as single-line objects.
When this constraint is not met, \thejoker\ will in general fail, and a model
that fits for a mixture of stellar spectra is more appropriate (e.g.,
\citealt{El-Badry:2018}).
The subset of RC stars are even more powerful as they are standard candles and
have masses that can be estimated using spectroscopy (using dredged-up elements;
\citealt{Martig:2016,Ness:2016}).
With mass estimates for the primary star, the binary-orbit fitting will return
$M_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
of the so-called ``binary mass function.''
Additionally, the \apogee\ pipelines (\citealt{Garcia-Perez:2016}) and also
\thecannon\ (\citealt{Ness:2015}) produce detailed abundance estimates for \RGB\
and \RC\ stars.
If there are causal relationships between chemical abundances and binary
companions---as are expected---these should be measurable.

By making cuts on this library of posterior \pdf\ samples (described in detail
in \sectionname~\ref{sec:catalogs}), we deliver catalogs of binary-star systems
from the \apogee\ survey, and show the bulk properties of these systems.

@@

\thejoker\ was designed for the extremely multi-modal \pdf 's expected when the
number of radial-velocity measurements of a source is small, or the data are
sparse (in phase-coverage) or noisy.
While other Markov Chain Monte Carlo (MCMC) methods have difficulty producing
independent samples with such data, \thejoker\ succeeds by brute force:
After generating an initial (very large) library of prior samples from an
assumed prior \pdf\ (see below), the (typically multi-modal) likelihood is
evaluated at each sample and used to rejection sample.
In practice, given a number of requested samples for each star, the sampling
proceeds iteratively: since it is easier to accept samples when the data is
sparse or noisy, far more prior sample draws (and thus likelihood evaluations)
must occur under very constraining data.

@@

In detail, we use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010})
implemented in \python\ (\package{emcee}; \citealt{Foreman-Mackey:2013}) to
perform the samplings.
We transform the standard Keplerian orbital parameters to a safer
parametrization, $(\ln P, \sqrt{K}\,\cos M_0, \sqrt{K}\,\sin M_0, \sqrt{e}\,\cos
\omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, for MCMC sampling.
This reparametrization is safer and more efficient for sampling with
\package{emcee}, which expects parameters to be components of a vector so that
linear operations can be applied (see, e.g., \citealt{Hogg:2017}); the angle
variables $(\omega, M_0)$ don't meet this requirement in the standard
parametrization.
We use the same prior \pdf 's as in \thejoker\ when running MCMC (see
\tablename~\ref{tbl:params}).

>>>>

The \apogee\ surveys primarily target red-giant-branch (\RGB) and other evolved
stars (e.g., red clump giants, RC), which are ideal for the study of single-line
binary systems.
In general, they are unlikely to have comparably-bright companions, and their
spectra are therefore fit as single-line objects.
When this constraint is not met, \thejoker\ will in general fail, and a model
that fits for a mixture of stellar spectra is more appropriate (e.g.,
\citealt{El-Badry:2018}).
The subset of RC stars are even more powerful as they are standard candles and
have masses that can be estimated using spectroscopy (using dredged-up elements;
\citealt{Martig:2016,Ness:2016}).
With mass estimates for the primary star, the binary-orbit fitting will return
$M_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
of the so-called ``binary mass function.''
Additionally, the \apogee\ pipelines (\citealt{Garcia-Perez:2016}) and also
\thecannon\ (\citealt{Ness:2015}) produce detailed abundance estimates for \RGB\
and \RC\ stars.
If there are causal relationships between chemical abundances and binary
companions---as are expected---these should be measurable.
By making cuts on this library of posterior \pdf\ samples (described in detail
in \sectionname~\ref{sec:catalogs}), we deliver catalogs of binary-star systems
from the \apogee\ survey, and show the bulk properties of these systems.

@@

\thejoker\ was designed for the extremely multi-modal \pdf 's expected when the
number of radial-velocity measurements of a source is small, or the data are
sparse (in phase-coverage) or noisy.
While other Markov Chain Monte Carlo (MCMC) methods have difficulty producing
independent samples with such data, \thejoker\ succeeds by brute force:
After generating an initial (very large) library of prior samples from an
assumed prior \pdf\ (see below), the (typically multi-modal) likelihood is
evaluated at each sample and used to reject all prior samples above the
likelihood surface (the rejection sampling step).
In practice, given a number of requested samples for each star, the sampling
proceeds iteratively: since it is easier to accept samples when the data is
sparse or noisy, far more prior sample draws (and thus likelihood evaluations)
must occur under very constraining data.

@@

In detail, we use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010})
implemented in \python\ (\package{emcee}; \citealt{Foreman-Mackey:2013}) to
perform the samplings.
\package{emcee} uses an ensemble of ``walkers'' (separate Markov chains) that
naturally adapt to the geometry of the parameter-space by generating proposal
steps based on the locations of other walkers (i.e. the ensemble chains are not
independent).
We transform the standard Keplerian orbital parameters to a safer
parametrization, $(\ln P, \sqrt{K}\,\cos (M_0 + \omega), \sqrt{K}\,\sin (M_0 +
\omega), \sqrt{e}\,\cos \omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, for MCMC
sampling.
This reparametrization is safer and more efficient for sampling with
\package{emcee}, which expects parameters to be components of a vector so that
linear operations can be applied (see, e.g., \citealt{Hogg:2017}); the angle
variables $(\omega, M_0)$ don't meet this requirement in the standard
parametrization.
We use the same prior \pdf 's as in \thejoker\ when running MCMC (see
\tablename~\ref{tbl:params}).
